Discrete Quantum Theories and Computing

### 2019-01-04 dissertation.tex
dissertation.lyx "Chapter 1 Introduction" Updating...

\centering{}\centering{}%
\begin{tabularx}{\linewidth}{ccYY}
\toprule 
\addlinespace
Physical quantities & Discrete & Discrete with extreme small unit & Continuous, or no evidence to support they are discrete\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational devices & Digital computer & Analog computer & Analog computer\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Mathematical representation & Integers & Real numbers & Real numbers\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational model  & Turing machine & BCSS machine & BCSS machine\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
How the models realize? & Reliably & 
Not reliably, but after we have a better technology to manipulate the minimum units, we can predict the physical phenomenon and their computational power by a better discrete model. & 
Not reliably, because precision can never be high enough, and the difference between their computational power might not be able compensated by error analysis techniques. \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Inspired quantum models &  & Quantum theories and computing over finite fields & Quantum interval-valued probability measures (QIVPM)\tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
The situations for classical quantities and the computational power
above them are summarized in Table~\ref{tab:Comparison}, and the
two possibilities dealing with real-valued models inspires us two
types of quantum models. Similar to classical analog computers, the
quantum circuit model, the most widely used model of quantum computing,
manipulates quantities which are not discrete apparently: the probability
amplitudes of quantum states. To address the similar inconsistency,
we first review the ``conventional quantum theory'' (CQT)\footnote{Alternative terminology in the literature includes ``actual,'' ``standard,''
and ``ordinary'' quantum theory.} in Chapter~\ref{chap:Conventional-Quantum-Theory}, where the probability
amplitudes are in the field of complex numbers. If the probability
amplitudes are assumed to be ultimately discrete, since their field
structure is essential to construct a quantum theory, we might assume
they are elements in discrete finite fields and build the quantum
theories and computing upon these finite fields in Chapter~\ref{chap:Unrestricted-Finite-Fields}
and \ref{chap:DQT=000026DQC}. Alternatively, the probability amplitudes
might never be found quantized so that we should try another approach
to take into account the precision of quantities. However, instead
of the precision of probability amplitudes, it is better to consider
the precision of their inducing probabilities because the probability
amplitudes is unlikely to be accessed directly while the probability
of an event can be deduced after a series of experiments. Moreover,
the idea of ``imprecise probabilities'' have well-studied classically
\cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
In Chapter~\ref{chap:QIVPM}, they will be extended it to quantum
domain, and define quantum interval-valued probability measures (QIVPM)
.

@@

Given a pure state~$\ket{\Phi}$, when we measure an observable represented
by an Hermitian matrix~$\mathbf{O}$, the measurement result is one
of the eigenvalues of $\mathbf{O}$. The probability of getting a
particular eigenvalue~$\lambda$ is $\melem{\Phi}{P}{\Phi}$, where
$P$ is the projection operator onto the eigenspace of $\lambda$.
This rule of computing the probability is called the Born rule \cite{Born1983,Mermin2007,Jaeger2007},
which is used when we want to extract information from a quantum computer.
For any mixed state~$\rho$, the generalized Born rule induces a
conventional quantum probability measure $\muB_{\rho}:\events\rightarrow[0,1]$,
where $\events$ is the set of all projection operators on a given
Hilbert space. Conversely, any quantum probability measure $\mu:\events\rightarrow[0,1]$
can be induced from a mixed state $\rho$ in the Hilbert space of
dimension $D\ge3$ according to Gleason's theorem \cite{gleason1957,Redhead1987-REDINA,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
In another word, this state $\rho$ is the unique state consistent
with any given quantum probability measure.
\end{comment}
The geometrical structure of states in CQT is first reviewed in Sec.~\ref{CQC1qubitBloch.sec},
where a pure state is represented as a vector in a Hilbert space,
and an irreducible state is actually a point in a projective Hilbert
space \cite{MosseriDandoloff2001,Jaeger2007} also known as a complex
projective space \cite{Hatcher2001,Bengtsson2007}. In Sec.~\ref{sec:fuzzy},
we will construct a quantum probability space based on classical probability
spaces, apply Gleason's theorem \cite{gleason1957,Redhead1987-REDINA,peres1995quantum,RichmanBridges1999,Hamhalter2013}
to recover the Born rule \cite{Born1983,Mermin2007,Jaeger2007}, and
define the expectation value of an observable. The expectation value
will then be used in Sec.~\ref{sec:The-geometry-of} to defined purity
which provides an easy way to decide whether a state is entangled
or not and to defined the maximum entangled states.

@@

\begin{comment}
Next, in Section \ref{discretequantumtheoryI}, we improve on this
by showing that for finite fields of order $p^{2}$, with the prime
$p$ of the form $4\ell+3$ ($\ell$ a non-negative integer), the
complex numbers have extremely compelling and natural discrete analogs
that permit a great many of the standard requirements of quantum computing
to be preserved. Under suitable conditions, we have amplitude-based
partitions of unity, unitary transformations, and entanglement, as
well as solutions to deterministic quantum algorithms such as the
algorithms of Deutsch, Simon, and Bernstein-Vazirani~\cite{NCbook,Mermin},
though still with some bothersome shortcomings. Because of the modular
nature of arithmetic in the finite complex field, it is not possible
to define an inner product in the usual sense, and we show in Section~\ref{discretequantumcomputingI}
that this leads to excessive computational power for the unstructured
database search problem for certain database sizes.
\end{comment}
\begin{comment}
Our second model called discrete quantum theory (I) considers only
finite fields of order $p^{2}$, with the prime $p$ of the form $4\ell+3$
($\ell$ a non-negative integer) \cite{geometry2013,DQT2014}. In
this model, an irreducible state is a vector in $\mathbb{F}_{p^{2}}^{2^{n}}$,
and can be reduced to a point in the discrete complex projective space~$\mathbb{DCP}^{2^{n}-1}$.
Among these irreducible states, the product states and entangled states
can not only be identified as in CQT but also be counted for different
$p$. Although the state space is a natural discrete analog to CQT,
this model still only predicts whether a measurement result is possible
or impossible, because it is not possible to define an inner product
in the usual sense due to the modular arithmetic~\cite{grove2002classical}.
Under suitable conditions, we can have deterministic quantum algorithms
such as the algorithms of Deutsch, Simon \cite{Simon:1994:PQC:1398518.1399019,Mermin2007,Jaeger2007},
and Bernstein-Vazirani \cite{Bernstein:1993:QCT:167088.167097,Mermin2007},
but this still leads to excessive computational power for the unstructured
database search problem for certain database sizes.
\end{comment}
In Sec.~\ref{discretequantumtheoryI} to \ref{discretequantumcomputingI},
we improve on this by showing that for finite fields of order $p^{2}$,
$\Fpp$, with the prime $p$ of the form $4\ell+3$ ($\ell$ a non-negative
integer), the complex numbers have extremely compelling and natural
discrete analogs that permit a great many of the standard requirements
of quantum computing to be preserved. In this model, the states of
a discrete $d$-dimensional system is a vector in the $d$-dimensional
vector space over $\Fpp$, and can be reduced to a point in the discrete
complex projective space. Among these irreducible states, the product
states and entangled states could not only be identified as in CQT
but also be counted for different $p$. Despite this model's similarity
to CQT, it still only predicts whether a measurement result is possible
or impossible, because it is not possible to define an inner product
in the usual sense due to the modular arithmetic~\cite{grove2002classical}.
Under suitable conditions, we can have deterministic quantum algorithms
such as the algorithms of Deutsch, Simon \cite{Simon:1994:PQC:1398518.1399019,Mermin2007,Jaeger2007},
and Bernstein-Vazirani \cite{Bernstein:1993:QCT:167088.167097,Mermin2007},
but this still leads to excessive computational power for partially
solving the $\usat$ problem efficiently.

@@

Finally, in Section~\ref{discretequantumcomputingII}, we apply our
discrete quantum theory to the study of two representative algorithms,
the deterministic Deutsch-Jozsa algorithm and the probabilistic Grover
algorithm~\cite{NCbook,Mermin}. The first algorithm highlights the
role played by the size of the field $p$ in determining the actual
resources required for computation as the number of input bits $n$
increases, a concept nonexistent in conventional quantum computing.
The second algorithm highlights, in addition, the dependence of the
precision of measurement (via cardinal probabilities) on the size
of the field, another nonexistent concept in conventional quantum
computing.
\end{comment}
\begin{comment}
Our third model, discrete quantum theory (II)~\cite{DQT2014}, restricts
states in some local region of $\mathbb{F}_{p^{2}}^{2^{n}}$. Within
a local region, a notion of inner product and probability could be
recovered. Its discrete quantum computing can be further applied to
the deterministic Deutsch-Jozsa algorithm \cite{DeutschJozsa1992,Jaeger2007}
and the probabilistic Grover algorithm \cite{Grover:1996:FQM:237814.237866,Mermin2007,Jaeger2007}.
\end{comment}
Since many important quantum algorithms are probabilistic, we want
to find a notion of probability for quantum theories over finite fields.
The cardinal probability defined in Sec.~\ref{sec:Discrete-Quantum-Theory-(II)}
restricts states in some local regions of the vector spaces. Within
a local region, a notion of inner product and probability could be
recovered. In Sec.~\ref{discretequantumcomputingII}, this discrete
quantum theory is applied on two representative algorithms, the deterministic
Deutsch-Jozsa algorithm \cite{DeutschJozsa1992,Jaeger2007} and the
probabilistic Grover algorithm \cite{Grover:1996:FQM:237814.237866,Mermin2007,Jaeger2007}.
\begin{comment}
When people tried to define quantum probability over finite fields,
people tended to treat the original Born rule as an axiom, and tried
to modify it to get a discrete Born rule \cite{Schumacher2012-SCHMQT,doi:10.1142/S0217984913500644,DQT2014,Ellerman2016a}.
However, any modified Born rule could hardly work on the whole vector
space, since there is no inner product on the whole vector space over
finite fields. Instead of treating the Born rule as an axiom, the
Born rule can actually be deduced from a set of abstract definitions
and axioms according to Gleason's theorem. Although we might hope
to deduce a discrete Born rule directly from a similar set of definitions
and axioms, no discrete Born rule satisfies certain properties motivated
by Gleason's theorem with infinitely precise real-number probability~\cite{Gardiner2014}.
Since the state spaces are now discrete and finite, this suggests
us to consider a discrete Born rule mapping to finite number of intervals
called interval-valued probability~\cite{JamisonLodwick2004,THOS2017}.
To adopting the idea of interval-valued probability step-by-step,
before attempting to study quantum interval-valued probability over
finite fields, we will first review the classical interval-valued
probability, and extend it with the conventional quantum theory.
\end{comment}
Despite the great success of cardinal probabilities, it is difficult
to define their arithmetic operations and expectation values. Since
these could be defined easily with real-valued probabilities in CQT,
we attempt to consider quantum probability measures over finite fields
(QPMFF) in Sec.~\ref{sec:Toward-IVPM}. In contrast to CQT, our Gleason-like
conditions over finite fields cannot deduce a discrete Born rule in
general. This impossible result suggests us that the discrete and
finite state spaces might not compatible with infinitely precise real-valued
probability. Thus, no matter whether the probability amplitudes are
ultimately discrete or not,

it might be more natural to consider a discrete Born rule mapping
to finite number of intervals, and we will start the discussion in
Chapter~\ref{chap:QIVPM}.

@@

The quantum extension, quantum interval-valued probability measure
(QIVPM) $\bar{\mu}:\events\rightarrow\mathscr{I}$, is a generalization
of both classical IVPMs $\bar{\mu}:2^{\Omega}\rightarrow\mathscr{I}$
and conventional quantum probability measures $\mu:\events\rightarrow[0,1]$
\cite{THOS2017}, because QIVPMs reduce to classical IVPMs when the
space of quantum events $\events$ is restricted to mutually commuting
events, and reduces to conventional quantum probability measures when
mapping to infinitely precise uncountable intervals $\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$.
While Shapley and Gleason both proved there must be a ``state''
consistent with any given QIVPM in the reduced cases, in general there
exists a QIVPM such that no state is consistent with it. However,
we found a class of QIVPMs such that all QIVPMs in this class are
consistent with a non-empty ``ball'' of quantum states whose radius
is defined by the maximal length of the intervals, and recovers the
original Gleason theorem asymptotically. Similarly, the conventional
quantum expectation value and the classical Choquet integral are together
generalized to the quantum interval-valued expectation value. This
is used to prove an imprecise Kochen-Specker theorem \cite{BELL_1966,kochenspecker1967,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
which suggests a possible resolution of the Meyer-Mermin debate on
the impact of finite-precision measurement on the Kochen-Specker theorem
\cite{PhysRevLett.83.3751,Mermin1999}.
\end{comment}
\begin{comment}
We present a mathematical framework based on quantum interval-valued
probability measures to study the effect of experimental imperfections
and finite precision measurements on defining aspects of quantum mechanics
such as contextuality and the Born rule. While foundational results
such as the Kochen-Specker and Gleason theorems are valid in the context
of infinite precision, they fail to hold in general in a world with
limited resources. Here we employ an interval-valued framework to
establish bounds on the validity of those theorems in realistic experimental
environments. In this way, not only can we quantify the idea of finite-precision
measurement within our theory, but we can also suggest a possible
resolution of the Meyer-Mermin debate on the impact of finite-precision
measurement on the Kochen-Specker theorem.
\end{comment}
Sec.~\ref{sec:Interval-Uncertainty} starts from reviewing how to
formulate classical ``imprecise probabilities'' \cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}
as interval-valued probability measures (IVPMs). Given a classical
IVPM, its core is the set of the real-valued probability measures
consistent with it \cite{Shapley1971,GilboaSchmeidler1994,NgMoYeh1997Chinese,Grabisch2016},
and the expectation value is \todo{Introduction Cleaned up until
here. }

Finally, we shift our attention to consider quantum interval-valued
probability measures (IVPMs), which potentially embody both finite
precision measurement and a sensible correspondence to standard quantum
probability. This interval-valued framework not only provides a natural
generalization of both classical IVPMs and conventional quantum probability
measures, but also allows us to establish bounds on the validity of
the Kochen-Specker and Gleason theorems in realistic experimental
environments.

\begin{sidewaystable}
\caption{Comparison among our models, where the upper half of the table lists
their quantum mechanical definitions and properties, and the lower
half of the table lists their computational power.}

\begin{tabularx}{\linewidth}{YYYYYYY}
\toprule 
\addlinespace
 & Conventional & Modal & Discrete (I) & Discrete (II) & QPMFF & QIVPM\tabularnewline\addlinespace
\addlinespace
\midrule
\midrule 
\addlinespace
 & Chapter~\ref{chap:Conventional-Quantum-Theory} & Sec.~\ref{modalquantum} & Sec.~\ref{discretequantumtheoryI} to \ref{sec:Geometry-of-Entangled-States} & Sec.~\ref{sec:Discrete-Quantum-Theory-(II)} & Sec.~\ref{sec:Toward-IVPM} & Chapter~\ref{chap:QIVPM}\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
States space & $\mathbb{C}^{D}$ & $\Fq^{D}$ & $\Fpp^{D}$ & Local region in $\Fpp^{D}$ & $\Fpp^{D}$ & $\mathbb{C}^{D}$\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Likelihood of events is predicted by & Real-valued probability & Possible or impossible & Possible or impossible & Cardinal probability & Real-valued probability, but no sensible Born rule & Interval-valued probability\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Expectation value & Defined & Undefined & Formally defined & Undefined &  & Defined\tabularnewline\addlinespace
\addlinespace
\midrule
\midrule 
\addlinespace
 &  & Sec.~\ref{modalquantumcomputing} & Sec.~\ref{discretequantumcomputingI} & Sec.~\ref{discretequantumcomputingII} &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Deutsch's algorithm & Yes & Maybe no & Yes & Yes &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Efficiently solve $\usat$ & Unlikely & Yes & Partially & Unlikely &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Grover search algorithm & Yes &  &  & Yes &  & \tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}

>>>>

\centering{}\centering{}%
\begin{tabularx}{\linewidth}{ccYY}
\toprule 
\addlinespace
Physical quantities & Discrete & Discrete with extreme small unit & Continuous, or no evidence to support they are discrete\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational devices & Digital computer & Analog computer & Analog computer\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Mathematical representation & Integers & Real numbers & Real numbers\tabularnewline\addlinespace
\addlinespace
\addlinespace
Computational model  & Turing machine & BCSS machine & BCSS machine\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
How the models realize? & Reliably & 
Not reliably, but after we have a better technology to manipulate the minimum units, we can predict the physical phenomenon and their computational power by a better discrete model. & 
Not reliably, because precision can never be high enough, and the difference between their computational power might not be able compensated by error analysis techniques. \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Inspired quantum models &  & Quantum theories and computing over finite fields & Quantum interval-valued probability measures (QIVPM)\tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
The situations for classical quantities and the computational power
above them are summarized in Table~\ref{tab:Comparison}, and the
two possibilities dealing with real-valued models inspires us two
types of quantum models. Similar to classical analog computers, the
quantum circuit model, the most widely used model of quantum computing,
manipulates quantities which are not discrete apparently: the probability
amplitudes of quantum states. To address the similar inconsistency,
we first review the ``conventional quantum theory'' (CQT)\footnote{Alternative terminology in the literature includes ``actual,'' ``standard,''
and ``ordinary'' quantum theory.} in Chapter~\ref{chap:Conventional-Quantum-Theory}, where the probability
amplitudes are in the field of complex numbers. If the probability
amplitudes are assumed to be ultimately discrete, we need to replace
the field of complex numbers by some discrete structures. Finite fields
is chosen because CQT is built upon the operations among the probability
amplitudes, and the probability amplitudes have the required operations
since the complex numbers form a field. The quantum theories and computing
built upon finite fields are discussed in Chapter~\ref{chap:Unrestricted-Finite-Fields}
and \ref{chap:DQT=000026DQC}. Alternatively, the probability amplitudes
might never be found quantized. In this situation, a better model
could take into account the precision of quantities. However, instead
of the precision of probability amplitudes, it is better to consider
the precision of their inducing probabilities because the idea of
``imprecise probabilities'' have well-studied classically \cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}.
In Chapter~\ref{chap:QIVPM}, they will be extended it to quantum
domain, and define quantum interval-valued probability measures (QIVPM).

@@

Given a pure state~$\ket{\Phi}$, when we measure an observable represented
by an Hermitian matrix~$\mathbf{O}$, the measurement result is one
of the eigenvalues of $\mathbf{O}$. The probability of getting a
particular eigenvalue~$\lambda$ is $\melem{\Phi}{P}{\Phi}$, where
$P$ is the projection operator onto the eigenspace of $\lambda$.
This rule of computing the probability is called the Born rule \cite{Born1983,Mermin2007,Jaeger2007},
which is used when we want to extract information from a quantum computer.
For any mixed state~$\rho$, the generalized Born rule induces a
conventional quantum probability measure $\muB_{\rho}:\events\rightarrow[0,1]$,
where $\events$ is the set of all projection operators on a given
Hilbert space. Conversely, any quantum probability measure $\mu:\events\rightarrow[0,1]$
can be induced from a mixed state $\rho$ in the Hilbert space of
dimension $D\ge3$ according to Gleason's theorem \cite{gleason1957,Redhead1987-REDINA,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
In another word, this state $\rho$ is the unique state consistent
with any given quantum probability measure.
\end{comment}
The geometrical structure of states in CQT is first reviewed in Sec.~\ref{CQC1qubitBloch.sec},
where a pure state is represented as a vector in a Hilbert space,
and an irreducible state is actually a point in a projective Hilbert
space \cite{MosseriDandoloff2001,Jaeger2007} also known as a complex
projective space \cite{Hatcher2001,Bengtsson2007}. In Sec.~\ref{sec:fuzzy},
we will construct a quantum probability space based on classical probability
spaces, apply Gleason's theorem \cite{gleason1957,Redhead1987-REDINA,peres1995quantum,RichmanBridges1999,Hamhalter2013}
to recover the Born rule \cite{Born1983,Mermin2007,Jaeger2007}, and
define the expectation value of an observable. The expectation value
will then be used in Sec.~\ref{sec:The-geometry-of} to define purity
which provides an easy way to decide whether a state is entangled
or not and to define the maximum entangled states.

@@

\begin{comment}
Next, in Section \ref{discretequantumtheoryI}, we improve on this
by showing that for finite fields of order $p^{2}$, with the prime
$p$ of the form $4\ell+3$ ($\ell$ a non-negative integer), the
complex numbers have extremely compelling and natural discrete analogs
that permit a great many of the standard requirements of quantum computing
to be preserved. Under suitable conditions, we have amplitude-based
partitions of unity, unitary transformations, and entanglement, as
well as solutions to deterministic quantum algorithms such as the
algorithms of Deutsch, Simon, and Bernstein-Vazirani~\cite{NCbook,Mermin},
though still with some bothersome shortcomings. Because of the modular
nature of arithmetic in the finite complex field, it is not possible
to define an inner product in the usual sense, and we show in Section~\ref{discretequantumcomputingI}
that this leads to excessive computational power for the unstructured
database search problem for certain database sizes.
\end{comment}
\begin{comment}
Our second model called discrete quantum theory (I) considers only
finite fields of order $p^{2}$, with the prime $p$ of the form $4\ell+3$
($\ell$ a non-negative integer) \cite{geometry2013,DQT2014}. In
this model, an irreducible state is a vector in $\mathbb{F}_{p^{2}}^{2^{n}}$,
and can be reduced to a point in the discrete complex projective space~$\mathbb{DCP}^{2^{n}-1}$.
Among these irreducible states, the product states and entangled states
can not only be identified as in CQT but also be counted for different
$p$. Although the state space is a natural discrete analog to CQT,
this model still only predicts whether a measurement result is possible
or impossible, because it is not possible to define an inner product
in the usual sense due to the modular arithmetic~\cite{grove2002classical}.
Under suitable conditions, we can have deterministic quantum algorithms
such as the algorithms of Deutsch, Simon \cite{Simon:1994:PQC:1398518.1399019,Mermin2007,Jaeger2007},
and Bernstein-Vazirani \cite{Bernstein:1993:QCT:167088.167097,Mermin2007},
but this still leads to excessive computational power for the unstructured
database search problem for certain database sizes.
\end{comment}
In Sec.~\ref{discretequantumtheoryI} to \ref{discretequantumcomputingI},
we improve on this by showing that for finite fields of order $p^{2}$,
$\Fpp$, with the prime $p$ of the form $4\ell+3$ ($\ell$ a non-negative
integer), the complex numbers have extremely compelling and natural
discrete analogs that preserve a great many of the standard requirements
of quantum computing. In this model, the states of a discrete $d$-dimensional
system is a vector in the $d$-dimensional vector space over $\Fpp$,
and can be reduced to a point in the discrete complex projective space.
Among these irreducible states, the product states and entangled states
could not only be identified as in CQT but also be counted for different
$p$. Despite this model's similarity to CQT, it still only predicts
whether a measurement result is possible or impossible, because it
is not possible to define an inner product in the usual sense due
to the modular arithmetic~\cite{grove2002classical}. Under suitable
conditions, we can have deterministic quantum algorithms such as the
algorithms of Deutsch, Simon \cite{Simon:1994:PQC:1398518.1399019,Mermin2007,Jaeger2007},
and Bernstein-Vazirani \cite{Bernstein:1993:QCT:167088.167097,Mermin2007},
but this still leads to excessive computational power for partially
solving the $\usat$ problem efficiently.

@@

Finally, in Section~\ref{discretequantumcomputingII}, we apply our
discrete quantum theory to the study of two representative algorithms,
the deterministic Deutsch-Jozsa algorithm and the probabilistic Grover
algorithm~\cite{NCbook,Mermin}. The first algorithm highlights the
role played by the size of the field $p$ in determining the actual
resources required for computation as the number of input bits $n$
increases, a concept nonexistent in conventional quantum computing.
The second algorithm highlights, in addition, the dependence of the
precision of measurement (via cardinal probabilities) on the size
of the field, another nonexistent concept in conventional quantum
computing.
\end{comment}
\begin{comment}
Our third model, discrete quantum theory (II)~\cite{DQT2014}, restricts
states in some local region of $\mathbb{F}_{p^{2}}^{2^{n}}$. Within
a local region, a notion of inner product and probability could be
recovered. Its discrete quantum computing can be further applied to
the deterministic Deutsch-Jozsa algorithm \cite{DeutschJozsa1992,Jaeger2007}
and the probabilistic Grover algorithm \cite{Grover:1996:FQM:237814.237866,Mermin2007,Jaeger2007}.
\end{comment}
Since many important quantum algorithms are probabilistic, we want
to find a notion of probability for quantum theories over finite fields.
The cardinal probability defined in Sec.~\ref{sec:Discrete-Quantum-Theory-(II)}
restricts states in some local regions of the vector spaces, where
a notion of inner product and probability could be recovered. In Sec.~\ref{discretequantumcomputingII},
this discrete quantum theory is applied on two representative algorithms,
the deterministic Deutsch-Jozsa algorithm \cite{DeutschJozsa1992,Jaeger2007}
and the probabilistic Grover algorithm \cite{Grover:1996:FQM:237814.237866,Mermin2007,Jaeger2007}.
\begin{comment}
When people tried to define quantum probability over finite fields,
people tended to treat the original Born rule as an axiom, and tried
to modify it to get a discrete Born rule \cite{Schumacher2012-SCHMQT,doi:10.1142/S0217984913500644,DQT2014,Ellerman2016a}.
However, any modified Born rule could hardly work on the whole vector
space, since there is no inner product on the whole vector space over
finite fields. Instead of treating the Born rule as an axiom, the
Born rule can actually be deduced from a set of abstract definitions
and axioms according to Gleason's theorem. Although we might hope
to deduce a discrete Born rule directly from a similar set of definitions
and axioms, no discrete Born rule satisfies certain properties motivated
by Gleason's theorem with infinitely precise real-number probability~\cite{Gardiner2014}.
Since the state spaces are now discrete and finite, this suggests
us to consider a discrete Born rule mapping to finite number of intervals
called interval-valued probability~\cite{JamisonLodwick2004,THOS2017}.
To adopting the idea of interval-valued probability step-by-step,
before attempting to study quantum interval-valued probability over
finite fields, we will first review the classical interval-valued
probability, and extend it with the conventional quantum theory.
\end{comment}
Despite the great success of cardinal probabilities, it is difficult
to define their arithmetic operations and expectation values. Since
these could be defined easily with real-valued probabilities in CQT,
we attempt to consider real-valued quantum probability measures over
finite fields (QPMFF) in Sec.~\ref{sec:Toward-IVPM}. In contrast
to CQT, our Gleason-like conditions over finite fields cannot deduce
a discrete Born rule in general. This impossible result suggests us
that the discrete and finite state spaces might not be compatible
with infinitely precise real-valued probability. Thus, no matter whether
the probability amplitudes are ultimately discrete or continuous,
we should study imprecise quantum probabilities formulated by quantum
interval-valued probability measures.

@@

The quantum extension, quantum interval-valued probability measure
(QIVPM) $\bar{\mu}:\events\rightarrow\mathscr{I}$, is a generalization
of both classical IVPMs $\bar{\mu}:2^{\Omega}\rightarrow\mathscr{I}$
and conventional quantum probability measures $\mu:\events\rightarrow[0,1]$
\cite{THOS2017}, because QIVPMs reduce to classical IVPMs when the
space of quantum events $\events$ is restricted to mutually commuting
events, and reduces to conventional quantum probability measures when
mapping to infinitely precise uncountable intervals $\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$.
While Shapley and Gleason both proved there must be a ``state''
consistent with any given QIVPM in the reduced cases, in general there
exists a QIVPM such that no state is consistent with it. However,
we found a class of QIVPMs such that all QIVPMs in this class are
consistent with a non-empty ``ball'' of quantum states whose radius
is defined by the maximal length of the intervals, and recovers the
original Gleason theorem asymptotically. Similarly, the conventional
quantum expectation value and the classical Choquet integral are together
generalized to the quantum interval-valued expectation value. This
is used to prove an imprecise Kochen-Specker theorem \cite{BELL_1966,kochenspecker1967,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
which suggests a possible resolution of the Meyer-Mermin debate on
the impact of finite-precision measurement on the Kochen-Specker theorem
\cite{PhysRevLett.83.3751,Mermin1999}.
\end{comment}
\begin{comment}
We present a mathematical framework based on quantum interval-valued
probability measures to study the effect of experimental imperfections
and finite precision measurements on defining aspects of quantum mechanics
such as contextuality and the Born rule. While foundational results
such as the Kochen-Specker and Gleason theorems are valid in the context
of infinite precision, they fail to hold in general in a world with
limited resources. Here we employ an interval-valued framework to
establish bounds on the validity of those theorems in realistic experimental
environments. In this way, not only can we quantify the idea of finite-precision
measurement within our theory, but we can also suggest a possible
resolution of the Meyer-Mermin debate on the impact of finite-precision
measurement on the Kochen-Specker theorem.
\end{comment}
Sec.~\ref{sec:Interval-Uncertainty} starts from reviewing how to
formulate classical ``imprecise probabilities'' \cite{Shafer1976,GilboaSchmeidler1994,Marinacci1999,JamisonLodwick2004,HuberRonchetti2009,Grabisch2016}
as interval-valued probability measures (IVPMs). Given a classical
IVPM, its core is the set of the real-valued probability measures
consistent with it \cite{Shapley1971,GilboaSchmeidler1994,NgMoYeh1997Chinese,Grabisch2016},
and the expectation value with respect to it is just classical Choquet
integral \cite{Choquet1954,GilboaSchmeidler1994,Grabisch2016}. When
we extend classical IVPMs to QIVPMs, these known properties of classical
IVPMs all have their natural correspondence for QIVPMs. In Sec.~\ref{sec:Kochen-Specker},
their expectation values will play a role in proving the Kochen-Specker
theorem \cite{BELL_1966,kochenspecker1967,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
for ``almost'' certain QIVPMs. In Sec.~\ref{sec:Gleason}, we are
able to show that, while a QIVPM incorporating the effects of finite
precision might not be consistent with Gleason's unique state~$\rho$
on all projectors defined on a Hilbert space~$\Hilb$ of dimension
$d\ge3$, there is a mathematically precise sense in which one recovers
the original Gleason theorem asymptotically. Specifically, it is possible
to construct a class of QIVPMs representing bounded resources that
is parameterized by the size of the intervals.

\begin{sidewaystable}
\caption{\label{tab:organization}Comparison among our models, where the upper
half of the table lists their quantum mechanical definitions and properties,
and the lower half of the table lists their computational power.}

\begin{tabularx}{\linewidth}{YYYYYYY}
\toprule 
\addlinespace
 & Conventional & Modal & Discrete (I) & Discrete (II) & QPMFF & QIVPM\tabularnewline\addlinespace
\addlinespace
\midrule
\midrule 
\addlinespace
 & Chapter~\ref{chap:Conventional-Quantum-Theory} & Sec.~\ref{modalquantum} & Sec.~\ref{discretequantumtheoryI} to \ref{sec:Geometry-of-Entangled-States} & Sec.~\ref{sec:Discrete-Quantum-Theory-(II)} & Sec.~\ref{sec:Toward-IVPM} & Chapter~\ref{chap:QIVPM}\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
States space & $\mathbb{C}^{D}$ & $\Fq^{D}$ & $\Fpp^{D}$ & Local region in $\Fpp^{D}$ & $\Fpp^{D}$ & $\mathbb{C}^{D}$\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Likelihood of events is predicted by & Real-valued probability & Possible or impossible & Possible or impossible & Cardinal probability & Real-valued probability, but no sensible Born rule & Interval-valued probability\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Expectation value & Defined & Undefined & Formally defined & Undefined &  & Defined\tabularnewline\addlinespace
\addlinespace
\midrule
\midrule 
\addlinespace
 &  & Sec.~\ref{modalquantumcomputing} & Sec.~\ref{discretequantumcomputingI} & Sec.~\ref{discretequantumcomputingII} &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Deutsch's algorithm & Yes & Maybe no & Yes & Yes &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Efficiently solve $\usat$ & Unlikely & Yes & Partially & Unlikely &  & \tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Grover search algorithm & Yes &  &  & Yes &  & \tabularnewline\addlinespace
\addlinespace
\bottomrule
\end{tabularx}
\end{sidewaystable}
The organization of this thesis is summarized in Table~\ref{tab:organization},
and we present our conclusions in final Chapter~\ref{chap:Conclusion}.
